index

### 2024-03-18 appendix/index-definition.qmd
dcor2d_2 index definition

-

@@

-   Note

  - From the function manual: "For bivariate data only, these are fast O(n log n) implementations of distance correlation and distance covariance statistics."

>>>>

    -   define pair-wised distance: $a_{ij} = \|y_1^i  - y_1^j\|$ and $b_{ij} = \|y_2^i  - y_2^j\|$ for $i, j = 1, 2, \ldots, n$.

    -   define the column sum, row sum, and overall sum: $$a_{i \cdot} = \sum_{l = 1}^n a_{il}, \quad a_{\cdot j} = \sum_{k = 1}^n a_{kj}, \quad a_{\cdot \cdot} = \sum_{k,l = 1}^n a_{kl} \quad b_{i \cdot} = \sum_{l = 1}^n b_{il}, \quad b_{\cdot j} = \sum_{k = 1}^n b_{kj}, \quad b_{\cdot \cdot} = \sum_{k,l = 1}^n b_{kl}$$

    -   define the distance covariance and variance: $$\text{dCov}(a_{ij}, b_{ij}) = \frac{1}{n(n-3)} \sum_{i \neq j} a_{ij} b_{ij} -\frac{2}{n (n-2)(n-3)}\sum_{i = 1}^n a_{i \cdot} b_{i \cdot} + \frac{1}{n(n - 1)(n-2)(n-3)} a_{\cdot \cdot} b_{\cdot \cdot}$$

    -   define the distance correlation: $$\text{dCor} (a_{ij}, b_{ij}) = \frac{\text{dCov}(a_{ij}, b_{ij})}{\sqrt{\text{dVar}(a_{ij}) \text{dVar}(b_{ij})}}$$

@@

    -   Xiaoming Huo &Gábor J. Székely (2016) Fast Computing for Distance Covariance: [link](https://www.tandfonline.com/doi/full/10.1080/00401706.2015.1054435). The papre proves that the covariance formula above (eq 3.3 in the paper) is an unbaised estimator of the population covariance in the original formula: $\displaystyle \operatorname {dCov} _{n}^{2}(X,Y):={\frac {1}{n^{2}}}\sum _{j=1}^{n}\sum _{k=1}^{n}A_{j,k}\,B_{j,k}$. The fast algorithm realises the new formula is in $O(n \log n)$ time.
    -   Maria L. Rizzo, Gábor J. Székely (2015) Energy distance: [link](https://wires.onlinelibrary.wiley.com/doi/abs/10.1002/wics.1375) WIREs Computational Statistics review paper
    -   Gábor J. Székely, Maria L. Rizzo, Nail K. Bakirov (2007) Measuring and testing dependence by correlation of distances: [link](https://projecteuclid.org/journals/annals-of-statistics/volume-35/issue-6/Measuring-and-testing-dependence-by-correlation-of-distances/10.1214/009053607000000505.full): original paper of distance correlation
